---
title: "Data Product week 4 Asignment"
author: "Sarah R"
date: "01/10/2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Human Development data

The United Nations Development Programme publishes data on Human Development Index (HDI) and key variables for all countries in the world. Data is available to download from the website <http://hdr.undp.org/en>. For this app, we use the 2015 figures for:

        1. Human development Index Ranking
        2. Education Index
        3. Gross Domestic Product(GDP) per capita (2011 PPP $)
        4. Carbon Dioxide Emissions per capita (tonnes)

```{r ind, echo = F}
path <- "/Users/srieubland/Documents/Documents/DataScience/09_DataProduct/week4/Asignment/"
dfEduc <- read.csv(paste0(path,"Education Index.csv"), stringsAsFactors = F, header = T, skip = 1)
dfGDP <- read.csv(paste0(path,"GDP.csv"), stringsAsFactors = F, header = T, skip = 1)
dfUnemp <- read.csv(paste0(path,"Total unemployment rate (% of labour force).csv"), stringsAsFactors = F, header = T, skip = 1)
dfEnvi <- read.csv(paste0(path,"Carbon dioxide emissions per capita (tonnes).csv"), stringsAsFactors = F, header = T, skip = 1)
```

```{r ind2, echo = F}
colnames(dfEduc)[1] <- "HDI_Rank"
dfIndex <- dfEduc[,c("HDI_Rank", "Country", "X2015")]
colnames(dfIndex)[3] <- "EducationIndex"

dfIndex$GDP_percapita <- dfGDP[match(dfIndex$Country, dfGDP$Country),]$X2015
dfIndex$Unemp_rate <- dfUnemp[match(dfIndex$Country, dfUnemp$Country),]$X2015
dfIndex$CO2_emi <- dfEnvi[match(dfIndex$Country, dfEnvi$Country),]$X2013

dfIndex[46,]$Country <- "  Cote d'Ivoire"
dfIndex$Country <- substr(dfIndex$Country,2, nchar(dfIndex$Country))
```

In order to plot the countries on the map, we download data on the average latitude and longitude for each country, from the opendata on Socrata.com (<https://opendata.socrata.com/>)

```{r coord, echo = F, cache = T}
fileUrl2 <- "https://opendata.socrata.com/api/views/mnkm-8ram/rows.csv?accessType=DOWNLOAD"
download.file(fileUrl2,destfile = "coordinates.csv",method = "curl")
dateDownloaded <- date()
```


```{r coord2, echo = F}
dfCoord <- read.csv("coordinates.csv", stringsAsFactors = F)
dfIndex$latitude <- dfCoord[match(dfIndex$Country, dfCoord$Country),]$Latitude..average.
dfIndex$longitude <- dfCoord[match(dfIndex$Country, dfCoord$Country),]$Longitude..average.
```

## Countries by HDI ranking (2015)

```{r map, echo = F, message= F}
dfIndex$name <- paste0(dfIndex$Country, " - Ranking = ", dfIndex$HDI_Rank)
dfIndex2 <- dfIndex[!is.na(dfIndex$lat), c("name", "latitude", "longitude")]

library(leaflet)
dfIndex2 %>% 
        leaflet() %>%
        addTiles() %>%
        addMarkers(popup = dfIndex2$name)
```

## App with R Shiny
My App  <https://sarahrieubland.shinyapps.io/app3/> allows the user to select countries by HDI ranking and visualise the linear model for key country variables.

### Education Index and GDP per capita
```{r model1, echo = F, message= F}
lm1 <- lm(GDP_percapita ~ EducationIndex , data = dfIndex)
plot(dfIndex$EducationIndex , dfIndex$GDP_percapita,
     ylab = "GDP per capita", xlab = "Education Index",
     #main = "Relationship between",
     cex = 1, pch = 16, bty = "n")
abline(lm1, col = "red", lwd = 2)
```

## CO2 Emissions per capita and GDP per capita
```{r model2, echo = F, message= F}
lm2 <- lm(GDP_percapita ~ CO2_emi , data = dfIndex)
plot(dfIndex$CO2_emi , dfIndex$GDP_percapita,
     ylab = "GDP per capita", xlab = "CO2 emission per capita [tonnes]",
     #main = "Relationship between",
     cex = 1, pch = 16, bty = "n")
abline(lm2, col = "blue", lwd = 2)
#points(lm1$model$GDP_percapita, lm1$model$CO2_emi, col = "red")
```
